---
title: "MSBD 5004 - Week 1-2"
author: "zzq"
date: "2017.02.16"
output:
  prettydoc::html_pretty:
    css: style.css
    highlight: null
    theme: hpstr
    toc: yes
runtime: shiny
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(shiny)
library(viridis)
library(ggthemes)
library(faraway)
#library(plotly)


```

\  
\  
\  


---

---

## 1. Fourier Series
  
**Fourier analysis** : Simple [sinusoids](https://en.wikipedia.org/wiki/Sine_wave) are the building blocks of the most complicated wave forms

\  
\  
\  

---  


###  1.1. Periodicity 
A function f(t) is periodic of period $T$ if there is a number $T>0$ such that
$$f(t+T) = f(t),$$

Is the sum of two periodic functions periodic? For instance, the following plot will show us the 
condition of $f(t) = 2sin(a*t) + cos(b*t)$

```{r, echo = FALSE}


s = seq(-10,10,0.01)

fluidPage(
    fluidRow(
        column(6,
            sliderInput("a", label = "a Input",
                        min = -10, max = 10, value = 10, step = 0.01 )
            ),
        column(6,    
            sliderInput("b", label = "b Input",
                        min = -10, max = 10, value = 9.5, step = 0.01 )
        )
    )    
)

s_active <- reactive({
    2*sin(input$a*s)+cos(input$b*s)
})

plot_data <- reactive({
    data.frame(list(s = s, s_active = s_active()))
})

renderPlot({
    ggplot(data = plot_data(),aes(#color = s,
                                  x = s, y = s_active)) +
        geom_line() +
        #scale_color_viridis(name = "t") +
        xlab("t") +
        ylab("f(t)")
})


```

In this example, $f$ will be a periodic function only if $a/b$ is a rational fraction $(b \not = 0)$.   

\  
\  
\  

---  

### 1.2. Building blocks

Considering about the sinusoid functions: $$Asin(2\pi\nu+\phi),$$

where $A$ is the Amplitude, $\nu$ is the frequency and $\phi$ is the phase.

> **Theorem 1.1**: Any periodic function can be approximated by sinusoid.  

>For instance:

>$$cos(t) = sin(t + \frac{\pi}{2})$$  



Noted that the amplitude term can be avoided:  

$$
\begin{align}
s(t) &= Asin(2\pi\nu t + \phi)  \\
     &= A[sin(2\pi\nu t)cos\phi + cos(2\pi\nu t)sin\phi] \\ 
     &= asin(2\pi\nu t) + bcos(2\pi\nu t)  
\end{align}
$$  

\  
\  
\  

---


### 1.3. Foundation of Fourier Analysis


> **Thorem 1.2** Let f(t) be **1-periodic** and $$\int_{0}^{1}|f(t)|^2 \leq \infty, $$ 
>then

>$$f(t) = \frac{a_0}{2} + \sum_{n=1}^\infty[a_n\cos(2\pi nt)+b_n\sin(2\pi nt)] $$

To calculate $a_n$ or $b_n$, we can calculate the $\int_0^1f(t)\cos(2\pi nt)dt$ or $\int_0^1f(t)\sin(2\pi nt)dt$. We can get following result based on the [orthogonality](https://en.wikipedia.org/wiki/Orthogonality#Orthogonal_functions) of this group of building blocks:

$$ \left\{
\begin{aligned}
a_n & =  2\int_0^1f(t)\cos(2\pi nt)dt \\
b_n & =  2\int_0^1f(t)\sin(2\pi nt)dt  \\
a_0 & =  2\int_0^1f(t)dt
\end{aligned}
\right.
$$


\  
\  
\  

---


### 1.4. Introducing complex exponentials

$$e^{i\theta} = \cos\theta + i\sin\theta $$

Here we can express *Theorem 1.2* by [complex exponentials](https://en.wikipedia.org/wiki/Euler%27s_formula): 

$$\begin{aligned}
\cos(2\pi nt) &= \frac{1}{2}(e^{2\pi int} + e^{2\pi i(-n)t}) \\
\sin(2\pi nt) &= \frac{1}{2i}(e^{2\pi int} + e^{2\pi i(-n)t}) \\ 
\end{aligned}
$$  

Finaly we have  

> **Theorem 1.3** Fourier Analysis
>$$f(t) = \sum_{n = -\infty}^{\infty}C_ne^{2\pi int}$$
  

$Cn$ can be calculated similiarly based on the orthogonality:  

$$C_n = \int_{0}^{1}f(t)e^{-2\pi int}dt$$

For a T-perioic function $f(t)$, g(t) = f(Tt) will have period 1. 

\  
\  
\  

---  

### 1.5. Some examples  

Find the Fourior series of the following periodic function: 
$$ f(t)=\left\{
\begin{aligned}
1&, N\leq t< N+1/2 \\
-1&, N-1/2 \leq t< N
\end{aligned}
\right., N \in Z
$$

```{r,echo=F}
eg1_t <- seq(-2.25,2.25,0.01)
eg1_y <- eg1_t %/% 0.5 %% 2 == 0 %>%
    as.numeric()
eg1_y[eg1_y==0] <- -1

eg1_data <- as.data.frame(cbind(eg1_t,eg1_y))

ggplot(data = eg1_data, aes(x = eg1_t, y = eg1_y)) +
    geom_line() +
    xlab("t") +
    ylab("f(t)")   

```  

This is a 1-Periodic function. Thus,
$$
\begin{aligned}
C_n &= \int_{0}^{1}f(t)e^{-2\pi int}dt \\
    &= \int_{0}^{1/2}f(t)e^{-2\pi int}dt + \int_{1/2}^{1}f(t)e^{-2\pi int}dt\\
    &= \frac{1}{-2\pi in}e^{-2\pi int}|^{1/2}_{0} + \frac{1}{2\pi in}e^{-2\pi int}|^{1}_{1/2}\\
    &= \frac{1}{-2\pi in}((-1)^n-1) + \frac{1}{2\pi in}(1-(-1)^n)\\
    &= \frac{1}{\pi in}(1-(-1)^n)\\
\end{aligned}
$$

Then we can write the Fourier series:  
$$
\begin{aligned}
f(t) &= \sum_{n\ odd} \frac{2}{\pi in}e^{2\pi int}, \ n = 2k+1 \\
     &= \sum_{k = -\infty}^{\infty}\frac{2}{\pi i(2k+1)}e^{2\pi i(2k+1)t}\\
     &= \sum_{k = 0}^{\infty}\frac{4}{\pi i(2k+1)}\sin{2\pi (2k+1)t}
\end{aligned}
$$

We can find something interesting if we let t=1/4:  

$$
1 - \frac{1}{3} + \frac{1}{5} - \frac{1}{7} + \frac{1}{9} - \frac{1}{11} + \cdots = \frac{\pi}{4}
$$

\  
\  
\  

---  


### 1.6. Gibbs phenomenon  

We can observate the [Gibbs Oscillation phenomenon](https://en.wikipedia.org/wiki/Gibbs_phenomenon) by using this example:  

```{r, echo = FALSE}


gibbs_t = seq(-1.5,1.5,0.01)

gibbs_y <- gibbs_t %/% 0.5 %% 2 == 0 %>%
    as.numeric()

gibbs_y[gibbs_y==0] <- -1

f_nt = function(k,t){
    4/(pi*(2*k+1))*sin(2*pi*(2*k+1)*t)
}

fluidPage(
    fluidRow(
        column(12,
            sliderInput("k", label = "k Input",
                        min = 0, max = 100, value = 10, step = 1 )
        )
    )    
)

gibbs_k <- reactive({
    0:input$k
})

gibbs_y_F <- reactive({
    gibbs_k() %>% 
        sapply(f_nt, gibbs_t) %>% 
        rowSums()
})

gibbs_data <- reactive({
    data.frame(list(t = gibbs_t,  y = gibbs_y, y_Fourier = gibbs_y_F() ))
})

renderPlot({
    ggplot(data = gibbs_data(),aes(#color = s,
                                  x = t, y = y)) +
        geom_line() +
        geom_line(aes(x = t, y = y_Fourier),color = "blue") +
        #scale_color_viridis(name = "t") +
        xlab("t") +
        ylab("f(t)") +
        theme_few() 
})


```  

> **The limitation of Fourier Analysis**: Not good to approximate a non-smooth function
 
The sum of the infinite Fourier series will be equal to the f(t). But it works not very well for a discontinuous function when the k is finite, since the sum of finite continuous functions is still continuous. 

\  
\  
\  

---


### 1.7. Some leftover stuff on Fourier Series

Let f(t) be i-periodic function, then we can define the energy:  

$$
\int_{0}^{1}|f(t)|^2dt \\  
f(t) = \sum_{n = -\infty}^{\infty}\hat{f}(n)e^{2\pi int}
$$

> **Theorem 1.4** $$\int_{0}^{1}|f(t)|^2dt = \sum_{n = -\infty}^{\infty}|\hat{f}(n)|^2 $$  
>Proof Hint:
$$
\hat{f}(n)\hat{f}(m)\int_{0}^{1}e^{2\pi i(n-m)t}dt = \left\{\begin{aligned}
1&, m = n \\
0&, m \not=n 
\end{aligned}
\right. 
$$

\  
\  
\  
\  
\  
\  

---  

---


## 2. Fourier Transform
Fourier transform is very similiar to Fourier series.  

> **Theorem 2.1** Fourier Transform can be defined as:
$$
\hat{f}(s) = \int_{-\infty}^{\infty}f(t)e^{-2\pi ist}dt
$$
> where s can be any real number. It transfroms a time domain function to spectral domain

The t、s can be regarded as "Time" and “Spectrum”  
Also the Inverse Fourier transform:
$$
\check{g}(t) = \int_{-\infty}^{\infty}g(s)e^{2\pi its}ds
$$
Note: For a symmetric function $f(t)$, we have $\check{\hat{f}} = f$

\  
\  
\  

---  

###  2.1. Some examples
#### Example 1

Find the Fourier transform of the following function:

$$
f(t)= \left\{\begin{aligned}
1&, -\frac{1}{2} < t < \frac{1}{2} \\
0&, otherwise 
\end{aligned}
\right. 
$$

The fourier transform process is shown as follows:

$$
\begin{aligned}
\hat{f}(s) &= \int_{-\infty}^{\infty}f(t)e^{-2\pi ist}dt = \int_{-\frac{1}{2}}^{\frac{1}{2}}1\cdot e^{-2\pi ist}dt \\
     &= \frac{e^{-2\pi ist}}{-2\pi is}\bracevert_{-\frac{1}{2}}^ {\frac{1}{2}}\\
     &= \frac{1}{-2\pi is}(e^{-\pi is} - e^{\pi is})\\
     &= \frac{sin(\pi s)}{\pi s}, \hspace{1cm} i.e. sinc\ function
\end{aligned}
$$

\  
\  
\  

#### Example 2  

Find the Fourier transform of the following function:

$$
f(t) = e^{-at^2}, \hspace{1cm} a > 0 
$$

we can find that:

$$
\hat{f}(s) = \sqrt{\frac{\pi}{a}}e^{-\frac{\pi^2}{a}s^2}
$$  
Noticed that if $a = \pi$, then $\hat{f}(s) = f(s)$.  
When $g(t)$ is the standarized normal distribution:  

$$
g(t) = \frac{1}{\sqrt{2\pi}\sigma}e^{-\frac{t^2}{2\sigma^2}}, \hspace{0.5cm} then \\
\hat{g}(s) = e^{-2\pi^2\sigma^2s^2}
$$

\  
\  
\  

---


### 2.2. Properties of Fourier transform 

In order to get a better understand of Fourier transform, we introduce $\mathcal{F}$ to denote Fourier transform:
$$
\mathcal{F}(f(t)) = \hat{f}(s)\\
\mathcal{F}^{-1}(g(s)) = \check{g}(t)
$$

1. Most important properties  
    + Linearity
    $$
    \mathcal{F}(f_1(t) + f_2(t)) = \mathcal{F}(f_1(t)) + \mathcal{F}(f_2(t))
    $$
    $$
    \mathcal{F}(af(t)) = a\mathcal{F}(f(t))
    $$
    
    + Invertibility
    $$
    \mathcal{F}^{-1}(\mathcal{F}(f(t))) = f(t)
    $$
    $$
    \mathcal{F}(\mathcal{F}^{-1}(g(s))) = g(s)
    $$
    
    + The value at 0s：
    $$
    \hat{f}(0) = \int_{-\infty}^{\infty}f(t)dt
    $$
    $$
    \check{g}(0) = \int_{-\infty}^{\infty}g(s)ds
    $$
    
    + Isometry
    $$
    \int_{-\infty}^{\infty}|f(t)|^2dt = 
    \int_{-\infty}^{\infty}|\hat{f}(s)|^2ds
    $$
    
2. Other useful properties
    + If $f(t)$ is real and symmetric, i.e. $f(-t) = f(t)$, then $\mathcal{F}(f(t))$ is a real function. On contrast, If $f(t)$ is real and anti-symmetric, i.e. $f(-t) = -f(t)$, then $\mathcal{F}(f(t))$ is imaginary.
    
    + Tricks on computation:
    $$
    \mathcal{F}(f(t+b)) = e^{2\pi isb}\hat{f}(s)
    $$
    $$
    \mathcal{F}(f(at)) = \frac{1}{a}\hat{f}(\frac{s}{a})
    $$
    $$
    \mathcal{F}(f(at+b)) = \frac{1}{a}e^{2\pi isb/a}\hat{f}(\frac{s}{a})
    $$

\  
\  
\  
    
---  

---  


## 3. Convolution
> Let $f(t), g(t)$ be functions, define the convolution process as: $$ (f*g)(t) = \int_{-\infty}^{\infty}f(y)g(t-y)dy $$

\  
\  
\  
  
---  


### 3.1. A baby example

Considering the moving average of f(t). The moving average can be described as:
$$
h(t) = \frac{1}{1}\int_{t-\frac{1}{2}}^{t+\frac{1}{2}}f(y)dy
$$

Here we define the moving window as indicator function: $g(t) = \textbf{1}_{[-\frac{1}{2},\frac{1}{2}]}(t)$, i.e.:
$$
g(t)= \left\{\begin{aligned}
1&, -\frac{1}{2} < t < \frac{1}{2} \\
0&, otherwise 
\end{aligned}
\right. 
$$

Then, 
$$
\begin{aligned}
h(t) &= \int_{-\infty}^{\infty}f(y)\cdot \textbf{1}_{[t-\frac{1}{2},t+\frac{1}{2}]}(y)dy\\
&= \int_{-\infty}^{\infty}f(y)g(t-y)dy
\end{aligned}
$$

\  

---  

### 3.2. Some properties of convolution  


1. $f*g = g*f$

+ Proof:
  $$
  (f*g)(t) = \int_{-\infty}^{\infty}f(y)g(t-y)dy
  $$
  let $z = t-y$, then $y = t-z$ and $dy = -dz$, then 
  $$
  \begin{aligned}
  (f*g)(t) &= \int_{\infty}^{-\infty}f(t-z)g(z)d(-z)\\
  &=\int_{-\infty}^{\infty}g(z)f(t-z)dz\\
  &=g*f
  \end{aligned}
  $$

2. $\mathcal{F}(f*g) = \mathcal{F}(f)\cdot\mathcal{F(g)}$

+ Proof:
  $$
  \begin{aligned}
  \mathcal{F}(f*g) &= \mathcal{F}(\int_{-\infty}^{\infty}f(y)g(t-y)dy)\\
  &= \int_{-\infty}^{\infty}e^{2\pi ist}(\int_{-\infty}^{\infty}f(y)g(t-y)dy)dt\\
  &= \int_{-\infty}^{\infty}f(y)(\int_{-\infty}^{\infty}e^{2\pi ist}g(t-y)dt)dy\\
  &= \int_{-\infty}^{\infty}f(y)(\hat{g}(s)e^{2\pi isy})dy\\
  &= \hat{g}(s)\hat{f}(s)
  \end{aligned}
  $$

\  

\  

---  

---  

**END**





